Design-based properties of some small-area estimators in forest inventory with two-phase sampling

### 2018-08-30 texfiles Daniel/techreporttwophase.tex
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

\begin{document}
\doublespacing
\nolinenumbers
\pagestyle{empty}

\title{Design-based properties of some small-area estimators in forest inventory with two-phase sampling \\ \small{revised version}}
\author{Daniel Mandallaz, Andreas Hill, Alexander Massey \\Deparment of Environmental Systems Science \\Chair of Land Use Engineering\\ETH Zurich\\CH 8092 Zurich, Switzerland}
\date{January 2016}

\maketitle
\newpage
$ $
\newpage
\begin{abstract}
We consider the small-area estimation problem in forest inventories with two-phase sampling schemes. We propose an improvement of the synthetic estimator, when the true mean of the auxiliary variables over the small-area is unknown and must be estimated, likewise for the residual corrected small-area estimator. We derive the asymptotic design-based variances of these new estimators, \textbf{the pseudo-synthetic and pseudo small-area estimators}, by incorporating also the design-based variance of the regression coefficients. We then propose a very simple mathematical device that transforms pseudo small-area estimators into pseudo-synthetic estimators, which is very convenient to derive asymptotic variances. The results are extended to cluster and two-stage sampling at the plot level. To illustrate the theory we consider the case of post-stratification and a case study.
\begin{center}
\textbf{R\'{e}sum\'{e}}
\end{center}
\begin{sloppypar}
Nous consid\`{e}rons le probl\`{e}me de l'estimation pour petits domaines dans le contexte d'inventaires forestiers en deux phases. Nous proposons une am\'{e}lioration simple de l'estimateur synth\'{e}tique quand la moyenne des variables auxiliaires dans le petit domaine doit \^{e}tre estim\'{e}e en premier lieu, de m\^{e}me pour l'estimateur pour petit domain bas\'{e} sur les r\'{e}sidus. Nous calculons la variance sous le plan de sondage de ces nouveaux estimateurs  en tenant compte de la variance des coefficients de r\'{e}gression.  De plus, nous proposons un artifice math\'{e}matique qui permet de transformer un estimateur pour petit domaine en un estimateur synth\'{e}tique, ce qui simplifie le calcul de la variance asymptotique. L'extension aux sondages par satellites et à deux degr\'{e}s au niveau de la placette est aussi trait\'{e}e. La th\'{e}orie est illustr\'{e}e par la post-stratification et par une \'{e}tude de cas.
\end{sloppypar}
\end{abstract}
\clearpage
\section*{Foreword}
The first version of this technical report has been published in the e-collection in September 2012. Since then many technical reports on further developments have been published in the e-collection
 (\cite{mandallazreport2,mandallazreport3,mandallazreport4,mandallazreport5,masseyphd}) as well as papers in the Canadian Journal of Forest Research, CJFR, (\cite{mandallaz3,mandallaz4,mandallaz5,massey1,massey2,massey3}. Recently, Dr. Alexander Massey and Andreas Hill, MSc, started writing an R-package implementing the various procedures discussed in the aforementioned publications.  Writing this software revealed some minor notational ambiguities used in cluster sampling as well as small numerical discrepancies between the new case study results produced in R and the original ones obtained with the IML procedure of the SAS package.  Detective work revealed that some SAS programs contained a minor error in a small-area variance formulae that caused the mean number of points per cluster in the small area to be taken from the second-phase instead of from the first-phase sample.  We emphasize that the underlying concept, i.e. "borrowing strength" by estimating regression coefficients over an unrestricted large area to improve the calculation of point and variance estimates in a restricted small area, remains precisely the same and that all the original conclusions of this technical report, including all relevant implications on subsequent works, remain valid.  This revised version improves the notation used for cluster sampling, presents updated tables with the corrected results for the case study and provides an enhanced discussion of the results. Some comments have been added to take into account results obtained since 2012. The mathematical derivations have also been carefully checked and, so far, confirmed.

@@

\pagestyle{plain}
\pagenumbering{arabic}

@@

Theoretically one may use the exact matrix $\pmb{A}$ if it is available or its estimate $\pmb{A}_{s_1}=\frac{1}{n_1}\sum_{x\in{s_1}}\pmb{Z}(x)\pmb{Z}^t(x)$ based on the large sample. However, the resulting point estimates are not always intuitively convincing and not optimal in the model-dependent framework. Besides, they are not available from the usual statistical software packages. For these reasons we shall only work with $\pmb{A}_{s_2}$.

@@

with the predictions $\hat{Y}_0(x)=\pmb{Z}^t(x)\pmb{\beta_0}$ and the residuals $R_0(x)=Y(x)-\hat{Y}_0(x)$, where $\pmb{\beta_0}$ is the given external regression coefficient, ideally obtained from another similar inventory whose sample points are independent of the current inventory's sample points. Note that in this case the mean residual will not necessarily be zero.

@@

where $\bar{\pmb{Z}}_G=\frac{1}{\lambda(G)}\int_G\pmb{Z}(x)dx$ is the true mean of the auxiliary vector over the small area $G$, which is available only if the first phase is exhaustive. $\hat{Y}_{G,synth}$ is unbiased under the model, but not optimal as it does not the take the model-dependent spatial correlation of the $Y(x)$ into account. Let us emphasize the fact that the model, i.e. $\hat{\pmb{\beta}}_{s_2}$, is fitted with the full data set and not only with $\{Y(x), \pmb{Z}(x)\mid x\in G\}$. \\
\textbf{In this paper we shall investigate the properties of $\hat{Y}_{G,synth}$ in the design-based inference framework}. \\
First, let us note that $\hat{Y}_{G,synth}$ is a design-based consistent sample copy of

@@

 Consequently, the synthetic estimator $\hat{Y}_{G,synth}$ has a design-based asymptotic bias equal to $-\frac{1}{\lambda(G)}\int_GR(x)$, which is not zero unless $G=F$ (we have zero mean residual over the entire domain, see [\ref{linearmodel2}]) or, which is unlikely, zero mean residual over the small area of interest. Using [\ref{synthetic1}] and [\ref{estvarmatrix}] \textbf{the estimated design-based variance of the synthetic estimator} is

@@

\hat{Y}_{G,small}=\hat{Y}_{G,synth}+

@@

which is clearly asymptotically equivalent to $\hat{Y}_{G,synth}$ as $n_1 \to \infty$ and its design-based expected value tends to  $\bar{\pmb{Z}}^t_G\pmb{\beta}$. To calculate the asymptotic variance we use the decomposition (actually the first order Taylor expansion)

@@

For very large $n_{1,G}$ it is clear that the external version will underestimate the variance

@@

For $G=F$ the $g^2_{F,s_1}(x)$ are asymptotically equal to $1$ (see \cite{mandallaz} p. 113 and the properties of the g-weights discussed below) so that both versions [\ref{varyregsynt1}] and [\ref{varexternalsimpleG3}] are asymptotically equivalent. However, version [\ref{varyregsynt1}] estimates the variance of the predictions in the large sample, which is better, and it rests upon the g-weights for the residual part, which is known to have better conditional properties (see \cite{mandallaz} p. 84 and section \ref{poststrat} for the important special case of stratification).\\

@@

\pmb{A}^{-1}_{s_2}\hat{\bar{\pmb{Z}}}_{2,G}}$.\\

@@

 By considering formally the trivial constant local density $Y(x)\equiv 1$ and solving the normal equations one sees that for any $G$ (i.e. also for $G=F$) one has

@@

  When $G=F$ the estimator $\hat{Y}_{reg}$ and the sample mean $\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)$ both converge towards the true value for an arbitrary $Y(x)$ and therefore one gets $\lim_{n_2\to\infty}g_{F,1}(x)=1>0$. This is not true for a proper subset $G\subset F$. In this case, $\hat{\tilde{Y}}_{G,synth}$ and the sample mean $\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}Y(x)$ converge towards the true value for an arbitrary $Y(x)$, which implies that $\tilde{g}_{G,1}(x)$ will tend to $0$ for $x\not\in{G}$ (negative values are possible) and to $\frac{n_2}{n_{2,G}} $ for $x\in{G}$.

@@

In simple random sampling one can derive non-parametric versions of the above procedures, i.e. kernel-based regression estimators or $k$ nearest neighbors estimators (\textbf{knn}). However, they are more difficult to implement than the regression estimators described above, primarily because on must use re-sampling (bootstrap) techniques to obtain reliable variance estimators because analytical results are beyond reach and the classical external variance formula severely underestimate the true variance as shown by simulations. Also, the performances of the non-parametric estimators are not better than the regression estimators (for global and small-area estimation) if the coefficient of determination $R^2$ is sufficient large, say above $0.7$, which is the case for timber volume assessment using modern remote sensing data as auxiliary information (e.g. canopy height parameters obtain via e.g. LiDAR). Details can be found in \cite{mandallazreport4}, \cite{massey2} and the PhD thesis \cite{masseyphd}.
In the next section we generalize the previous results on regression estimators to cluster sampling, widely used in national inventories. The main ideas remain the same but the formulae are slightly more cumbersome due to the random cluster size. The non-parametric estimators are not yet available for cluster sampling.

@@

We follow the description of cluster sampling as defined in \cite{mandallaz} (especially section 5.5). A cluster is identified by its origin $x$, uniformly distributed in $\tilde{F}\supset F$. The geometry of the cluster is given by $M$ vectors $e_1,\ldots e_M$ defining the random cluster $x_l=x+e_l$. $M(x)=\sum_{l=1}^MI_{F}(x_l)$ is the random number of points of the cluster falling into the forest area $F$. We define the local density at the cluster level by $Y_c(x)=\frac{\sum_{l=1}^MI_{F}(x_l)Y(x_l)}{M(x)}$, likewise we set $\pmb{Z}_c(x)=\frac{\sum_{l=1}^MI_{F}(x_l)\pmb{Z}(x_l)}{M(x)}$. The set $\tilde{F}$ above can be mathematically defined as the smallest set $\{x\in{\mathcal{R}}^2 \mid M(x) \ne 0 \}$. In the first phase we have $n_1$ clusters identified by $x\in{s_1}$ and in the second phase $n_2$ clusters with $x\in{s_2}$, obtained by simple random sampling from $s_1$.\\

@@

In the pure design-based approach the weights will be $M^2(x)$ but this leads to non-zero mean residual (though close zero in practice), and the definitions of the regression estimator and of the normal equation are slightly different (see \cite{mandallaz}, section 5.5 for details). The choice of $M(x)$ rather than $M^2(x)$ as weights is suggested by the model-dependent approach. When $Y_c(x)$ is the mean of the

@@

The small-area estimation set up is slightly more complicated because the points of a cluster with $M(x) \ge 0, \;x\in{\tilde{F}}$ can be spread over more than one small area. The notation used in the 2012 version of this technical report was not sufficiently explicit in this respect. This does not affect the regression coefficients, always based on the entire sample in $F$ (the borrowing strength philosophy), but the formulation "after restricting the samples to the small area of interest" is now more explicit in the notation. We define the following quantities:
\begin{eqnarray} \label{extranotation}
M_G(x) &=& \sum_{l=1}^L  I_G(x+e_l) \nonumber \\
Y_{c,G}(x)&=&\frac{\sum_{l=1}^L I_G(x+e_l)Y(x+e_l)}{M_G(x)} \nonumber \\
\hat{\bar{Y}}_{c,2,G} &=& \frac{\sum_{x\in{s_{2,G}}}M_G(x)Y_{c,G}(x)}{\sum_{x\in{s_{2,G}}}M_G(x)}\nonumber \\
\pmb{Z}_{c,G}(x)&=& \frac{\sum_{l=1}^L I_G(x+e_l)\pmb{Z}(x+e_l)}{M_G(x)} \nonumber \\
\hat{Y}_{c,G}(x)&=&\pmb{Z}^t_{c,G}(x)\hat{\pmb{\beta}}_{c,s_2} \nonumber \\
\hat{R}_{c,G}&=& Y_{c,G}(x)-\hat{Y}_{c,G}(x) \nonumber \\
\hat{\bar{\pmb{Z}}}_{c,1,G}&=&\frac{\sum_{x\in{s_{1,G}}}M_G(x)\pmb{Z}_{c,G}(x)}{\sum_{x\in{s_{1,G}}}M_G(x)}

@@

The estimated variance of the one-phase estimator is given by (\cite{mandallaz}, section 4.3).
\begin{equation}\label{onephaseclustervariance}
\hat{\VAR}(\hat{\bar{Y}}_{c,2,G})=\frac{1}{n_{2,G}(n_{2,G}-1))}
\sum_{x\in{s_{2,G}}}\big(\frac{M_G(x)}{\bar{M}_{2,G}}\big)^2(Y_{c,G}(x)-\hat{\bar{Y}}_{c,2,G})^2
\end{equation}
\noindent where $\bar{M}_{2,G}=\frac{\sum_{x\in{s_{2,G}}}M_G(x)}{n_{2,G}}$.\\
The estimate of the covariance matrix of the auxiliary variables is given by
\begin{equation}\label{estcovauxiliarycluster}
\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}_{c,1,G}}=\frac{1}{n_{1,G}(n_{1,G}-1)}
\sum_{x\in{s_{1,G}}}\big(\frac{M_G(x)}{\bar{M}_{1,G}}\big)^2(\pmb{Z}_{c,G}(x)-\hat{\bar{\pmb{Z}}}_{c,1,G})
(\pmb{Z}_{c,G}(x)-\hat{\bar{\pmb{Z}}}_{c,1,G})^t
\end{equation}
\noindent where $\bar{M}_{1,G}=\frac{\sum_{x\in{s_{1,G}}}M_G(x)}{n_{1,G}}$.\\
The \textbf{pseudo-synthetic estimate} is then
\begin{eqnarray}\label{psynthclusterest}
\hat{Y}_{c,G,psynth}&=&\hat{\bar{\pmb{Z}}}_{c,1,G}^t\hat{\pmb{\beta}}_{c,s_2}\nonumber \\
&=& \frac{1}{n_2}\sum_{x\in{s_2}}g_{c,1,G}(x)Y_c(x)
\end{eqnarray}
with the g-weights $g_{c,1,G}(x)=\hat{\bar{\pmb{Z}}}_{c,1,G}^t\pmb{A}^{-1}_{c,s_2}M(x)\pmb{Z}_c(x)$. The estimated variance is as in [\ref{estvarpseudosynth1}]

\begin{equation}\label{estvarpseudosynth1cluster}
\hat{\VAR}(\hat{Y}_{c,G,psynth})=
\hat{\bar{\pmb{Z}}}_{c,1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{c,s_2}}\hat{\bar{\pmb{Z}}}_{c,1,G}
+ \hat{\pmb{\beta}}_{c,s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{c,1,G}}\hat{\pmb{\beta}}_{c,s_2}
\end{equation}
The pseudo-synthetic estimate is generally design-biased. Adjusting for the residuals we get the  \textbf{pseudo small-area estimator}
\begin{equation}\label{psmallclusterest}
\hat{Y}_{c,G,psmall}=\hat{\bar{\pmb{Z}}}_{c,1,G}^t\hat{\pmb{\beta}}_{c,s_2}
+ \frac{\sum_{x\in{s_{2,G}}}M_G(x)\hat{R}_{c,G}(x)}{\sum_{x\in{s_{2,G}}}M_G(x)}
\end{equation}
It is asymptotically design-unbiased and, by analogy with simple random sample (see [\ref{pseudosmallvar}]) its variance can be expected to be well approximated by

\begin{eqnarray}\label{estvarpsmallcluster}
\hat{\VAR}(\hat{Y}_{c,G,psmall})&=&
\hat{\bar{\pmb{Z}}}_{c,1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{c,s_2}}\hat{\bar{\pmb{Z}}}_{c,1,G}
+ \hat{\pmb{\beta}}_{c,s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{c,1,G}}\hat{\pmb{\beta}}_{c,s_2}
\nonumber \\ &+& \frac{1}{n_{2,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}\big(\frac{M_G(x)}{\bar{M}_{2,G}}\big)^2
(\hat{R}_{c,G}(x)-\bar{\hat{R}}_{2,G})^2
\end{eqnarray}
where $\bar{M}_{2,G}=\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}M_G(x)$ and
$\bar{\hat{R}}_{2,G}=\frac{\sum_{x\in{s_{2,G}}}M_G(x)\hat{R}_{c,G}(x)}{\sum_{x\in{s_{2,G}}}M_G(x)}$.
One can rewrite [\ref{estvarpsmallcluster}] with g-weights and predictions as in [\ref{pseudosmallvar2}].
\par

@@

$I_{c,G}(x)=\frac{\sum_{l=1}^M I_G(x_l)}{M(x)}=\frac{M_G(x)}{M(x)}$. We shall assume that $M_G(x)\equiv M(x)$ for all $x\in{\tilde{G}}=\{x \mid \sum_{l=1}^M I_G(x_l)>0\}$. If we have a partition of the forest in small areas $G_k, k=1,2,\ldots L$, then $\tilde{F} \subset \cup_{k=1}^L \tilde{G}_k$ but it is unlikely in general that the $\tilde{G}_k$ are disjoined. We can reasonably hope that in extensive forest inventory the surface area of points where $M_G(x)< M(x)$ is negligible. The theoretical normal equation

@@

\hat{\tilde{Y}}_{c,G,psynth}=\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}^t\hat{\pmb{\theta}}_{c,s_2}

@@

$\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}=\frac{\sum_{x\in{s_{1,G}}}M_G(x)\pmb{\mathcal{Z}}_{c,G}(x)}
{\sum_{x\in{s_{1,G}}}M_G(x)}$

@@

@@

\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{c,s_2}}
\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}

@@

\sum_{x\in{s_{1,G}}}\big(\frac{M_G(x)}{\bar{M}_{1,G}}\big)^2(\pmb{\mathcal{Z}}_{c,G}(x)-

@@

(\pmb{\mathcal{Z}}_{c,G}(x)-\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G})^t

@@

\textbf{Remark}:\\
Even if the condition $M_G(x)\equiv M(x)$ is fulfilled in $s_{2,G}$ we cannot expect $\hat{\tilde{Y}}_{c,G,psynth}$ to be asymptotically unbiased (although the sample copy of the normal equation converges towards the theoretical normal equation this does not guarantee zero mean residuals over $G$ as we have seen). However the bias can be expected to be negligible.\\[0.5cm]
Obviously a decomposition similar to [\ref{extsynthethic3}] will hold so that $\hat{\tilde{Y}}_{c,psynth}$ and $\hat{Y}_{c,G,psmall}$ in [\ref{psmallclusterest},\ref{estimatorclusterextended}]
will be close to each other in large samples.\\

@@

\tilde{g}_{c,1,G}(x)=\hat{\bar{\pmb{\mathcal{Z}}}}^t_{c,1,G}\pmb{\mathcal{A}}^{-1}_{c,s_2}M(x)

@@

\hat{\tilde{Y}}_{c,G,psynth}=\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{c,1,G}(x)Y_c(x)

@@

\tilde{g}^2_{c,1,G}(x)\hat{\mathcal{R}}^2(x)+

@@

 It follows from general principles presented in (\cite{mandallaz}, sections 4.4 and 4.5) that one can use all the previous results by replacing everywhere the exact local densities $Y(x)$, or $Y(x_l)$ in cluster sampling, by the corresponding generalized local densities $Y^*(x)$ or $Y^*(x_l)$. The second-stage variance is automatically taken into account. More mathematical details are given in \cite{mandallazreport5}.

@@

We reanalyze the case study described with in Chapter 9 of \cite{mandallaz}. The inventoried area covered $218ha$.  The auxiliary information is based on 16 stands defined by the following qualitative variables obtained from the manual interpretation of aerial photographs:

>>>>

\begin{document}
\doublespacing
\pagestyle{plain}
\pagenumbering{arabic}

\title{Design-based properties of some small-area estimators in forest inventory with two-phase sampling}
\author{Daniel Mandallaz \thanks{Tel. ++41(0)44 6323186 e-mail
    daniel.mandallaz@env.ethz.ch} \\Chair of Land Use Engineering\\ETH Zurich\\CH 8092 Zurich, Switzerland}
\date{}

\maketitle

@@

\nolinenumbers
\begin{abstract}
We consider the small-area estimation problem in forest inventories with two-phase sampling schemes. We propose an improvement of the synthetic estimator, when the true mean of the auxiliary variables over the small-area is unknown and must be estimated, likewise for the residual corrected small-area estimator. We derive the asymptotic design-based variances of these new estimators, \textbf{the pseudo-synthetic and pseudo small-area estimators}, by incorporating also the design-based variance of the regression coefficients. We then propose a very simple mathematical device that transforms pseudo small-area estimators into pseudo-synthetic estimators, which is very convenient to derive asymptotic variances. The results are extended to cluster and two-stage sampling at the plot level. To illustrate the theory we consider the case of post-stratification and a case study.
\begin{center}
\textbf{R\'{e}sum\'{e}}
\end{center}
\begin{sloppypar}
Nous consid\`{e}rons le probl\`{e}me de l'estimation pour petits domaines dans le contexte d'inventaires forestiers en deux phases. Nous proposons une am\'{e}lioration simple de l'estimateur synth\'{e}tique quand la moyenne des variables auxiliaires dans le petit domaine doit \^{e}tre estim\'{e}e en premier lieu     , de m\^{e}me pour l'estimateur pour petit domain bas\'{e} sur les r\'{e}sidus. Nous calculons la variance sous le plan de sondage de ces nouveaux estimateurs  en tenant compte de la variance des coefficients de r\'{e}gression.  De plus, nous proposons un artifice math\'{e}matique qui permet de transformer un estimateur pour petit domaine en un estimateur synth\'{e}tique, ce qui simplifie le calcul de la variance asymptotique. L'extension aux sondages par satellites et à deux degr\'{e}s au niveau de la placette est aussi trait\'{e}e. La th\'{e}orie est illustr\'{e}e par la post-stratification et par une \'{e}tude de cas.
\end{sloppypar}
\end{abstract}
\clearpage

@@

Theoretically one may use the exact matrix $\pmb{A}$ if it is available or its estimate $\pmb{A}_{s_1}=\frac{1}{n_1}\sum_{x\in{s_1}}\pmb{Z}(x)\pmb{Z}^t(x)$ based on the large sample. However, the resulting point estimates are not always intuitively convincing and not optimal in the model-dependent framework. Beside, they are not available from the usual statistical software packages. For these reasons we shall only work with $\pmb{A}_{s_2}$.

@@

with the predictions $\hat{Y}_0(x)=\pmb{Z}^t(x)\pmb{\beta_0}$ and the residuals $R_0(x)=Y(x)-\hat{Y}_0(x)$, where $\pmb{\beta_0}$ is the given external regression coefficient, ideally obtained from another similar inventory. Note that in this case the mean residual will not necessarily be zero.

@@

where $\bar{\pmb{Z}}_G=\frac{1}{\lambda(G)}\int_G\pmb{Z}(x)dx$ is the true mean of the auxiliary vector over the small area $G$, which is available only if the first phase is exhaustive. $\hat{Y}_{G,synt}$ is unbiased under the model, but not optimal as it does not the take the model-dependent spatial correlation of the $Y(x)$ into account. Let us emphasize the fact that the model, i.e. $\hat{\pmb{\beta}}_{s_2}$, is fitted with the full data set and not only with $\{Y(x), \pmb{Z}(x)\mid x\in G\}$. \\
\textbf{In this paper we shall investigate the properties of $\hat{Y}_{G,synt}$ in the design-based inference framework}. \\
First, let us note that $\hat{Y}_{G,synt}$ is a design-based consistent sample copy of

@@

 Consequently, the synthetic estimator $\hat{Y}_{G,synt}$ has a design-based asymptotic bias equal to $-\frac{1}{\lambda(G)}\int_GR(x)$, which is not zero unless $G=F$ (we have zero mean residual over the entire domain, see [\ref{linearmodel2}]) or, which is unlikely, zero mean residual over the small area of interest. Using [\ref{synthetic1}] and [\ref{estvarmatrix}] \textbf{the estimated design-based variance of the synthetic estimator} is

@@

\hat{Y}_{G,small}=\hat{Y}_{G,synt}+

@@

which is clearly asymptotically equivalent to $\hat{Y}_{G,synt}$ as $n_1 \to \infty$ and its design-based expected value tends to  $\bar{\pmb{Z}}^t_G\pmb{\beta}$. To calculate the asymptotic variance we use the decomposition (actually the first order Taylor expansion)

@@

For very large $n_{1,G}$ is is clear that the external version will underestimate the variance

@@

For $G=F$ the $g^2_{F,s_1}(x)$ are asymptotically equal to $1$ (see \cite{mandallaz} p. 113 and the properties of the g-weights discussed below) so that both versions [\ref{varyregsynt1}] and [\ref{varexternalsimpleG3}] are asymptotically equivalent. However, version [\ref{varyregsynt1}] estimates the variance of the predictions in the large sample, which is better, and it rests upon the g-weights for the residual part, which it is known to have better conditional properties (see \cite{mandallaz} p. 84 and section \ref{poststrat} for the important special case of stratification).\\

@@

pmb{A}^{-1}_{s_2}\hat{\bar{\pmb{Z}}}_{2,G}}$.\\

@@

 By considering formally the trivial constant local density $Y(x)\equiv 1$ and solving the normal equations one sees that

@@

  When $G=F$ the estimator $\hat{Y}_{reg}$ is asymptotically equivalent to the sample mean over $F$, i.e. to $\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)$. This must hold for an arbitrary density $Y(x)$ and therefore one gets $\lim_{n_2\to\infty}g_{F,1}(x)=1>0$. This is not true for a proper subset $G\subset F$. In this case, $\hat{\tilde{Y}}_{G,synt}$ is asymptotically equivalent to the sample mean over the small domain, i.e. to $\hat{\bar{Y}}_G$. Hence, $\tilde{g}_{G,1}(x)$ will tend to $0$  for $x\not\in{G}$ (negative values are possible) and to $\frac{n_2}{n_{2,G}} $ for $x\in{G}$.

@@

In the next section we generalize the previous results to cluster sampling. The main ideas remain the same but the formulae are slightly more cumbersome due to the random cluster size.

@@

We follow the description of cluster sampling as define in \cite{mandallaz} (especially section 5.5). A cluster is identified by its origin $x$, uniformly distributed in $\tilde{F}\supset F$. The geometry of the cluster is given by $M$ vectors $e_1,\ldots e_M$ defining the random cluster $x_l=x+e_l$. $M(x)=\sum_{l=1}^MI_{F}(x_l)$ is the random number of points of the cluster falling into the forest area $F$. We define the local density at the cluster level by $Y_c(x)=\frac{\sum_{l=1}^MI_{F}(x_l)Y(x_l)}{M(x)}$, likewise we set $\pmb{Z}_c(x)=\frac{\sum_{l=1}^MI_{F}(x_l)\pmb{Z}(x_l)}{M(x)}$. The set $\tilde{F}$ above can be mathematically defined as the smallest set $\{x\in{\mathcal{R}}^2 \mid M(x) \ne 0 \}$. In the first phase we have $n_1$ clusters identified by $x\in{s_1}$ and in the second phase $n_2$ clusters with $x\in{s_2}$, obtained by simple random sampling from $s_1$.\\

@@

In the pure design-based approach the weights will be $M^2(x)$ but this leads to non-zero mean residual (thought close zero in practice), and the definitions of the regression estimator and of the normal equation are slightly different (see \cite{mandallaz}, section 5.5 for details). The choice of $M(x)$ rather than $M^2(x)$ as weights is suggested by the model-dependent approach. When $Y_c(x)$ is the mean of the

@@

In two-phase sampling we estimate the mean of the auxiliary information over the small area $G$ by
$$\hat{\bar{\pmb{Z}}}_{c,1,G}=\frac{\sum_{x\in{s_{1,G}}}M(x)\pmb{Z}_c(x)}{\sum_{x\in{S_{1,G}}}M(x)}$$
with estimated covariance matrix
\begin{equation}\label{estcovauxiliarycluster}
\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}_{c,1,G}}=\frac{1}{n_{1,G}(n_{1,G}-1)}
\sum_{x\in{s_{1,G}}}\big(\frac{M(x)}{\bar{M}_{1,G}}\big)^2(\pmb{Z}_c(x)-\hat{\bar{\pmb{Z}}}_{c,1,G})
(\pmb{Z}_c(x)-\hat{\bar{\pmb{Z}}}_{c,1,G})^t
according to (\cite{mandallaz}, section 4.3).
The \textbf{pseudo-synthetic estimate} is then
\begin{eqnarray}\label{psynthclusterest}
\hat{Y}_{c,G,psynth}&=&\hat{\bar{\pmb{Z}}}_{c,1,G}^t\hat{\pmb{\beta}}_{c,s_2}\nonumber \\
&=& \frac{1}{n_2}\sum_{x\in{s_2}}g_{c,1,G}(x)Y_c(x)

@@

with the g-weights $g_{c,1,G}(x)=\hat{\bar{\pmb{Z}}}_{c,1,G}^t\pmb{A}^{-1}_{c,s_2}M(x)\pmb{Z}_c(x)$. The estimated variance is as in  [\ref{estvarpseudosynth1}]

\begin{equation}\label{estvarpseudosynth1cluster}
\hat{\VAR}(\hat{Y}_{c,G,psynth})=
\hat{\bar{\pmb{Z}}}_{c,1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{c,s_2}}\hat{\bar{\pmb{Z}}}_{c,1,G}
+ \hat{\pmb{\beta}}_{c,s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{c,1,G}}\hat{\pmb{\beta}}_{c,s_2}
\end{equation}
The pseudo-synthetic estimate is generally design-biased. Adjusting for the residuals we get the  small-area estimator
\begin{equation}\label{psmallclusterest}
\hat{Y}_{c,G,psmall}=\hat{\bar{\pmb{Z}}}_{c,1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{c,s_2}}
+ \frac{\sum_{x\in{s_{2,G}}}M(x)\hat{R}_c(x)}{\sum_{x\in{s_{2,G}}}M(x)}
\end{equation}
It is asymptotically design-unbiased and intuitively its variance can be expected to be approximated by

\begin{eqnarray}\label{estvarpsmallcluster}
\hat{\VAR}(\hat{Y}_{c,G,psmall})&=&
\hat{\bar{\pmb{Z}}}_{c,1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{c,s_2}}\hat{\bar{\pmb{Z}}}_{c,1,G}
+ \hat{\pmb{\beta}}_{c,s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{c,1,G}}\hat{\pmb{\beta}}_{c,s_2}
\nonumber \\ &+& \frac{1}{n_{2,g}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}\big(\frac{M(x)}{\bar{M}_{2,G}}\big)^2
(\hat{R}_c(x)-\bar{\hat{R}}_{2,G})^2
where $\bar{M}_{2,G}=\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}M(x)$ and
$\bar{\hat{R}}_{2,G}=\frac{\sum_{x\in{s_{2,G}}}M(x)\hat{R}_c(x)}{\sum_{x\in{s_{2,G}}}M(x)}$.
One can rewrite [\ref{estvarpsmallcluster}] with g-weights and predictions as in [\ref{pseudosmallvar2}].
\par

@@

$I_{c,G}(x)=\frac{\sum_{l=1}^M I_G(x_l)}{M(x)}$. In extensive inventories we can reasonably assume that all the points of a cluster lying in the forest area $F$ will belong to the same small area $G$ so that in fact $I_{c,G}(x) \equiv 1$ for all $x\in{\tilde{G}}=\{x \mid \sum_{l=1}^M I_G(x_l)>0\}$. The theoretical normal equation

@@

\hat{\tilde{Y}}_{c,psynth}=\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}^t\hat{\pmb{\theta}}_{c,s_2}

@@

$\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}=\frac{\sum_{x\in{s_{1,G}}}M(x)\pmb{\mathcal{Z}}_c(x)}
{\sum_{x\in{s_{1,G}}}M(x)}$

@@

Obviously a decomposition similar to [\ref{extsynthethic3}] will hold so that $\hat{\tilde{Y}}_{c,psynth}$ and $\hat{Y}_{c,G,small}$ in [\ref{psmallclusterest},\ref{estimatorclusterextended}]
are asymptotically equivalent.\\

@@

\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{c,s_2}}\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}

@@

\sum_{x\in{s_{1,G}}}\big(\frac{M(x)}{\bar{M}_{1,G}}\big)^2(\pmb{\mathcal{Z}}_c(x)-

@@

(\pmb{\mathcal{Z}}_c(x)-\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G})^t

@@

@@

\tilde{g}_{G,c,1}(x)=\hat{\bar{\pmb{\mathcal{Z}}}}^t_{c,1,G}\pmb{\mathcal{A}}^{-1}_{c,s_2}M(x)

@@

\hat{\tilde{Y}}_{c,G,psynth}=\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{G,c,1}(x)Y_c(x)

@@

\tilde{g}^2_{G,c,1}(x)\hat{\mathcal{R}}^2(x)+

@@

 It follows from general principles presented in (\cite{mandallaz}, sections 4.4 and 4.5) that one can use all the previous results by replacing everywhere the exact local densities $Y(x)$, or $Y(x_l)$ in cluster sampling, by the corresponding generalized local densities $Y^*(x)$ or $Y^*(x_l)$. The second-stage variance is automatically taken into account.

@@

We reanalyze the case study described with full details in Chapter 9 of \cite{mandallaz}. The inventoried area covered $218ha$.  The auxiliary information is based on 16 stands defined by the following qualitative variables:
